---
title: "Homework 1"
author: "Bui Le Linh"
date: '2017-09-20'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1 (Chapter 2, #1, 8 marks)

(a) Flexible method is probably better. Since we have lots of data, flexible statistical learning method would achieve lower bias. Large n also helps reduce overfitting.

(b) Flexible method is probably worst. Flexible statistical learning method would overfit, leading to high variance. 

(c) Flexible method is probably better. Flexible statistical learning method would achieve lower bias by learning non-linear relationship among predictors. Inflexible statistical learning method would underfit and have high bias.

(d) Flexible method is probably worst. Flexible statistical learning method would "chase the error", leading to high variance.

## Question 2 (Chapter 2, #2, 6 marks)

(a) This is Regression problem because the dependent variable CEO salary is continious. We are interested in inference. n = 500, p = 3.

(b) This is Classification problem because the dependen variable is binary (success/failure). We are interested in prediction. n = 20, p = 13.

(c) This is Regression problem because the dependent variable % change in US dollar is continious. We are interested in prediction (and probably inference as well). n = 52, p = 3.

## Question 3 (Chapter 2, #9, 8 marks)

```{r}
library(ISLR) # use install.packages("ISLR") to install
data(Auto)
head(Auto)
```

(a) Variables:
```{r}
Auto <- na.omit(Auto)
summary(Auto)
```
"name" is qualitative. The remaining variables are all quantitative.
Please note that some variables such as origin are probably quanlitative, but have been encoded using numerical values. So, R treat these variables as quantitative, unless we use factor() to indicate a variable as qualitative/categorical.

(b) Range of quantitative variables:

```{r}
myr <- function(x) {
  out <- range(x)
  names(out) <- c("min","max")
  out
}
apply(Auto[,1:8],MARGIN=2,FUN=myr)
```


(c) Mean and Standard deviation of quantitative variables:

```{r}
meansd <- function(x) {
  out<-c(mean(x),sd(x))
  names(out) <- c("mean","SD")
  out
}
apply(Auto[,1:8],MARGIN=2,FUN=meansd)
```


(d) Range, Mean and Standard deviation of quantitative variables in the subset:

```{r}
AutoSubset <- Auto[-c(10:85),]
apply(AutoSubset[,1:8],MARGIN=2,FUN=myr)
apply(AutoSubset[,1:8],MARGIN=2,FUN=meansd)
```

(e) Relationship among predictors:
```{r}
pairs(Auto)
```
mpg seems to negatively correlated with displacement, horsepower, weight, and positively correlated with year.
cylinder seems to be positively correlated with displacement, horsepower, weight.
displacement seems to be positively correlated with horsepower, weight, and negatively correlated with acceleration.
horsepower seems to be positively correlated with weight and negtatively correlated with acceleration.

(f) Displacement, horsepower, weight and year would be an useful variable to predict mpg. Since the correlations between mpg and displacement, horsepower and weight are so similar, we should probably only need one of them in our model (instead of all three).
